\chapter*{AI Declaration}
\addcontentsline{toc}{chapter}{AI Declaration}

The usage of AI tools within this project is documented here. I solemnly declare that I have documented all interactions with AI tools, including the prompts used and the outputs received.

\renewcommand{\arraystretch}{1.5}
{\footnotesize
\begin{longtable}{p{0.20\textwidth} p{0.40\textwidth} p{0.40\textwidth}}
    \toprule
    System & Prompt & Usage \\
    \midrule
    \endfirsthead
    \toprule
    System & Prompt & Usage \\
    \midrule
    \endhead
    \midrule
    \endfoot
    \bottomrule
    \endlastfoot
    GitHub Copilot 1 & Asked to delete the existing project and provide a simple LaTeX template with a shared preamble, a Chapters subdirectory, TOC, glossary, abbreviations, and Roman numerals for non-content pages. & Template structure and LaTeX setup provided \\
    GitHub Copilot 2 & Requested additional title page lines (WAB header, reviewer, module) for the main document. & Title page metadata and layout updated \\
    GitHub Copilot 3 & Requested uppercase Roman numerals for front matter and lowercase Roman numerals for back matter page numbering. & Page numbering adjusted in main and exposee \\
    GitHub Copilot 4 & Requested exposee title page to include shared info fields (WAB header, department, module, reviewer). & Exposee title page updated to include shared metadata \\
    GitHub Copilot 5 & Requested adding the provadis-hochschule.pdf logo to the top right corner of both main and exposee title pages. & Logo added to top right of both title pages \\
    GitHub Copilot 6 & Reported that Glossary and Abbreviations sections were missing from compiled output. & Added example \texttt{\textbackslash newacronym} entries to abbreviations.tex\\
    GitHub Copilot 7 & Reported that glossary section still not appearing in final PDF. & Updated settings.tex to include \texttt{automake} option in glossaries package to enable automatic glossary generation \\
    GitHub Copilot 8 & Requested Python benchmark code for LCS comparison with multiple synthetic scenarios (random strings, implanted mutated substrings, etc.) & Created benchmark script and documentation in Code folder \\
    GitHub Copilot 9 & Asked to add a progress-bar-like output. & Added optional single-line in-place progress output with percentage, completed/total steps, elapsed time, ETA, and \texttt{--progress}/\texttt{--no-progress} flags \\
    GitHub Copilot 10 & Reported that ETA was inaccurate due to varying string lengths & Reworked ETA estimation to be algorithm- and length-aware using observed per-bucket runtimes\\
    GitHub Copilot 11 & Requested separate build and query timing (and plots). & Refactored benchmark to record build/query time separately and added total time summaries \\
    GitHub Copilot 12 & Requested plots using statistics beyond the mean (median, standard deviation, etc.). & Expanded summary stats and generated mean+std and median+IQR plots alongside memory plots \\
    GitHub Copilot 13 & Requested splitting benchmarking and plotting so plot generation can run separately, with a shared CLI entry point and default run-then-plot behavior. & Added CLI mode switch (run/plot/both) to reuse saved CSV results for plotting without rerunning benchmarks \\
    GitHub Copilot 14 & Requested an error if the two algorithms return different LCS result strings for the same input. & Added strict substring equality check (in addition to length) and raise an error on mismatch \\
    GitHub Copilot 15 & Requested that, in case of multiple valid LCS results, both algorithms return all results and compare the sets. & Updated both algorithms to return all max-length substrings and compare result sets for correctness \\
    GitHub Copilot 16 & Requested refactoring the benchmark script into helper modules under a helpers subdirectory, with separate algorithm imports and plotting helper, and clean top-level imports in the main file. & Split benchmark code into dedicated helper modules (SAM, ESA, plotting, generation, statistics, benchmarking) and simplified main script orchestration \\
    GitHub Copilot 17 & Requested mean, median, standard deviation, and IQR for memory usage plots as well. & Added memory quartile aggregation and generated memory mean+SD plus median+IQR plots \\
    GitHub Copilot 18 & Asked for a practical memory/space-complexity metric and whether build-time and query-time memory can be measured separately. & Added separate build/query phase memory metrics and a persistent index-size metric, with aggregation and plots for comparison \\
    GitHub Copilot 19 & Asked whether adding total time (build + query) columns and graphs with full statistics would make sense. & Added full total-time statistics (mean, median, std, IQR, Q1, Q3) to summaries and created dedicated total-time plots \\
    GitHub Copilot 20 & Requested moving plot legends away from the top center because they obscured titles. & Repositioned all figure legends to the right side outside the plotting area to keep titles unobstructed \\
    GitHub Copilot 21 & Requested an additional output file containing both generated strings and their LCS value(s). & Added export of one CSV per generated case with `s`, `t`, LCS length, and all LCS substrings \\
    GitHub Copilot 22 & Asked to replace the lambda with a tuple-key-based approach. & Replaced lambda-based sorting with a tuple-key-based approach for suffix array construction, improving performance \\
    GitHub Copilot 23 & Asked to refactor the plotting code to reduce repetition and make it more modular. & Refactored plotting code to use a single generic plotting function\\
    GitHub Copilot 24 & Asked to replace the suffix array construction with the SA-IS algorithm. & Replaced doubling-based suffix array construction with an SA-IS implementation (linear-time induced sorting with LMS recursion), keeping the same public API\\
    GitHub Copilot 25 & Reported that generated benchmark tables do not fit on the page and requested a smart code-side solution (splitting tables if needed) instead of shrinking them. & Updated the table-export generator to split each algorithm summary into multiple narrower tables (timing, memory, index size)\\
    GitHub Copilot 26 & Requested that the generated benchmark tables should not show up in the List of Tables. & Adjusted the generator to avoid captions and instead use a custom numbered title with \texttt{\textbackslash refstepcounter\{table\}} and \texttt{\textbackslash label} for cross-references\\
    GitHub Copilot 27 & Requested that the generated tables are still slightly too large and should use a smaller font size. & Wrapped generated tabular blocks in \texttt{\textbackslash small} within the generator to reduce width without making tables unreadably tiny\\
\end{longtable}
}
