\chapter*{Appendix}
\addcontentsline{toc}{chapter}{Appendix}

The full summary statistics (mean, standard deviation, median, \acrshort{iqr}, minimum, maximum, quartiles, and sample counts) per scenario, algorithm, and metric 
are available in the supplementary materials at \path{Code/Results/summary_stats.csv}.

The full raw benchmark data can be found at \path{Code/Results/raw_runs.csv}, and the corresponding string pairs with their LCS values are stored in \path{Code/Results/case_lcs_values.csv}.

% Write your appendix here.

\section*{Slope Ratios}
\label{app:sloperatios}
The following tables present the slope ratios between the \acrshort{sam} and the \acrshort{esa} for each scenario and metric, 
along with the coefficient of determination ($R^2$) as a measure of goodness of fit for the linear regression.

\begin{itemize}
    \item \textbf{SAM slope} and \textbf{ESA slope}: Fitted slope of the linear regression of the median measurement against string length for the respective algorithm. 
    The slope is to be interpreted as the constant $a$ in the linear model $y = a \cdot x + b$, where $y$ is the median measurement (e.g., build time) and $x$ is the length of one of the strings in the string pair.
    \item \textbf{SAM / ESA ratio}: Ratio between the two slopes. Values below $1$ mean that the \acrshort{sam} grows more slowly than the \acrshort{esa}, while values above $1$ mean the opposite.
    \item \textbf{SAM $R^2$} and \textbf{ESA $R^2$}: A measure of how closely a linear model fits the observed median values for each algorithm, with values closer to $1$ indicating a stronger linear relationship.
    \item \textbf{Factor}: An expression of how many times larger the greater slope is compared to the smaller one, always reported as a value $\geq 1$. 
    For example, a factor of $17.30\times$ means one algorithm's growth rate is 17.30 times steeper than the other's.
    \item \textbf{Worse}: The algorithm with the steeper slope. As in, the one that scales less favourably with increasing input length.
\end{itemize}

% Sorted and formatted benchmark tables
\subsection*{Build Time}
\input{Code/Results/tables/slope_ratios_build_time.tex}

\subsection*{Query Time}
\input{Code/Results/tables/slope_ratios_query_time.tex}

\subsection*{Total Time}
\input{Code/Results/tables/slope_ratios_total_time.tex}

\subsection*{Index Size}
\input{Code/Results/tables/slope_ratios_index_size.tex}

\subsection*{Peak Memory Usage}
\input{Code/Results/tables/slope_ratios_peak_memory.tex}

